Modelling the gas kinematics of an atypical Lyman alpha emitting compact dwarf galaxy

### 2017-06-25 paper/paper.tex
updated referee reply

\begin{abstract}
  The conditions in star-forming Compact Dwarf Galaxies (CDGs) 
  place them as the best systems to test models on primordial
galaxy formation and evolution.    
Here we report on one of such CDGs, \tol, which presents
a broad, single peaked, highly symmetric \lya\ line emission that had
evaded theoretical interpretation so far.  
In this paper we explain these features by two different physically
motivated kinematic models: 
an interstellar medium composed by outflowing clumps with 
random motions and an homogeneous gaseous sphere undergoing solid body
rotation.
It is the first time that an observed \lya\ spectrum can be explained
assuming either of these kinematic conditions.
The multiphase model requires a clump velocity dispersion of
\sigmaclump\ with outflows of \inftyclump\ while the bulk rotation
velocity is constrained to be $348^{+75}_{-48}$ \kms.
This shows the importance of considering multiphase
and rotation kinematics among the possible conditions shaping the
\lya\ spectra.
However, we consider that the multiphase model has more chances to
agree with \tol's reality. 
The clump velocity dispersion implies a dynamical mass of $2\times
10^{9}$ M$_{\odot}$, ten times its baryonic mass, which can be
explained if \tol\ is hosted by a dark matter halo of at least
$10^{11}$ M$_{\odot}$; conditions similar to those of Dwarf Elliptical
Galaxies. 
If future kinematic maps of \tol\ confirm the velocities found by the
multiphase model, it could provide a new perspective to study
evolutionary processes linking star-forming with quiescent phases in
primordial galaxy evolution. 
\end{abstract}

@@

We computed the projected half-light radius to be $R_s=1.5\pm0.1$ kpc 
from the surface brightness profiles reported by \cite{2003A&A...410..481N}. 
Assuming spherical geometry, one can translate this value into a 3D
half-light radius of $r_s=3R_s/2=2.25$ kpc.

@@

We divide the results in two classes: low $\chi^2/d.o.f.<35$ and high
$\chi^2/d.o.f.35>$.
We apply a random forest classification algorithm to find the
parameters more relevant in splitting the results into these two
classes. The algorithm uses $500$ trees for the classification and a
maximum of $3$ depth levels. 
To check for tability we repeat the computation $10$ times by
randomly subsampling the input data to use $80\%$ of the data as a
training set.

@@

Figure \ref{fig:tree} shows as an illustration the results for a
classification tree.  
The tree starts with $28$ and $1962$ models in the low and high
$\chi^2/d.o.f.$ classes, respectively. 
We find that the best classification
yields $13$ and $44$ models in the low and high $\chi^2\d.o.f.$
classes, respectively, after selecting for $v_{\infty,{\rm cl}}<157.0$ \kms,
$\sigma_{{\rm cl}}>55.6$ \kms and $P_{\rm cl}>0.683$. 
The split in two classes is not perfect. 
This means that the 3 parameters in the
model are the most influential, but that some combinations of the
other 11 parameters are also relevant to find a good fit.

@@

As a test, we run a small set of 100 additional radiative transfer simulations
with random values for all parameters, except for $\sigma_{\rm cl}$,
$v_{\infty {\rm cl}}$, $P_{\rm cl}$ which are within the derived
constraints.
Consistent with the results from the random forest algorithm,
illustrated by the classification tree of Figure \ref{fig:tree}, we
find that $30$ models show $\chi^2/d.o.f.<35$.

@@

We cannot discard that another region of parameter space with similar
velocity structure but an empty inter-clump medium, relatively dense
clumps and at least a few clumps per sightline 
could also work \citep{Hansen06}.
This possibility does not change our conclusions about the
kinematic parameters, but it might be interesting to explore in the
future with a higher sampling of the parameter space once new 
observations  clarify the kinematic nature of \tol \footnote{In
  Section \S \ref{sec:conclusions} we describe  in more detail what
  kind of observations could be performed in the near future to
  improve our understanding of \tol's kinematics.} .

@@

\begin{equation}
M_{\rm dyn} = 3 \frac{\sigma ^{2}r}{G} = 3.48\times10^{9}
\left(\frac{\sigma}{100\ \mathrm{km\ s}^{-1}}\right)^2\left(\frac{r}{\mathrm{kpc}}\right)
M_{\odot}, 
\end{equation}
%
Assuming that the \lya\ emission is entirely powered by star formation 
we use the 3D half-light radius $r_s=2.25$ kpc as the typical size
for the HI region, in the multiphase model the clump velocity dispersion
$\sigma_{\rm{cl}}=$\sigmaclump corresponds to a dynamical mass of
$M_{\rm dyn}=2.31\pm0.04 \times 10^{9}$ $M_{\odot}$, which is ten
times the estimated baryonic mass in \tol.

@@

We acknowledge that a final confirmation of our kinematic findings has
to be provided by new observational tests with integral field unit
spectroscopy. 
If our results are confirmed, this would open up a new line of more detailed
enquiries to understand how a compact dwarf galaxy can have the
structural characteristics similar to a dwarf elliptical while being
at the same time isolated and sustaining large starburst.
Such and investigation could provide an evolutionary link between star-forming
and quiescent stages of dwarf galaxy evolution.

>>>>

\begin{abstract}
  The conditions in star-forming Compact Dwarf Galaxies (CDGs) 
  place them as the best systems to test models on primordial
galaxy formation and evolution.    
Here we report on one of such CDGs, \tol, which presents
a broad, single peaked, highly symmetric \lya\ line emission that had
evaded theoretical interpretation so far.  
In this paper we explain these features by two different physically
motivated kinematic models: 
an interstellar medium composed by outflowing clumps with 
random motions and an homogeneous gaseous sphere undergoing solid body
rotation.
It is the first time that an observed \lya\ spectrum can be explained
assuming either of these kinematic conditions.
The multiphase model requires a clump velocity dispersion of
at least \sigmaclump\ with outflows of \inftyclump\ at most, while the
bulk rotation velocity is constrained to be $348^{+75}_{-48}$ \kms.
This shows the importance of considering multiphase
and rotation kinematics among the possible conditions shaping the
\lya\ spectra.
However, we consider that the multiphase model has more chances to
agree with \tol's reality. 
The clump velocity dispersion implies a dynamical mass of at least $2\times
10^{9}$ M$_{\odot}$, ten times its baryonic mass, which can be
explained if \tol\ is hosted by a dark matter halo of at least
$10^{11}$ M$_{\odot}$; conditions similar to those of Dwarf Elliptical
Galaxies. 
If future kinematic maps of \tol\ confirm the velocities found by the
multiphase model, it could provide a new perspective to study
evolutionary processes linking star-forming with quiescent phases in
primordial galaxy evolution. 
\end{abstract}

@@

We computed the projected half-light radius to be $R_s=1.5\pm0.1$ kpc 
from the surface brightness profiles reported by \cite{2003A&A...410..481N}. 
Assuming spherical geometry, one can translate this value into a 3D
half-light radius of $r_s=3R_s/2=2.25$ kpc.
Imaging observations by \cite{2001AJ....121..169F} show that \tol\ is
an isolated field galaxy not belonging to a group or cluster.

@@

We divide the results in two classes: low $\chi^2/d.o.f.<35$ and high
$\chi^2/d.o.f.35>$.
We apply a random forest classification algorithm to find the
more relevant parameters in splitting the results into these two
classes. 
The algorithm uses $500$ trees for the classification and a
maximum of $3$ depth levels. 
To check for tability we repeat the computation $10$ times by
randomly subsampling the input data to use $80\%$ of the data as a
training set.

@@

Figure \ref{fig:tree} shows as an illustration the results for a
classification tree.  
The tree starts with $28$ and $1962$ models in the low and high
$\chi^2/d.o.f.$ classes, respectively. 
We find that the best classification
yields $13$ and $44$ models in the low and high $\chi^2\d.o.f.$
classes, respectively, after selecting for $v_{\infty,{\rm cl}}<157.0$ \kms,
$\sigma_{{\rm cl}}>55.6$ \kms and $P_{\rm cl}>0.683$. 
The split in two classes is not perfect. 
This means that the 3 parameters in the
model are the most influential in finding models with low
$\chi^2/d.o.f.$, but that some combinations of the other 11 parameters
are also relevant to find a good fit.

@@

As a test, we also run a small set of 100 additional radiative
transfer simulations with random values for all parameters, except for
$\sigma_{\rm cl}$, $v_{\infty {\rm cl}}$, $P_{\rm cl}$ which are within the derived
constraints.
Consistent with the results from the random forest algorithm,
illustrated by the classification tree of Figure \ref{fig:tree}, we
find that $30$ models show $\chi^2/d.o.f.<35$.

@@

We cannot discard that another region of parameter space
could also work.
It might be interesting to explore in the
future with a higher sampling of the parameter space once new 
observations  clarify the kinematic nature of \tol \footnote{In
  Section \S \ref{sec:conclusions} we describe  in more detail what
  kind of observations could be performed in the near future to
  improve our understanding of \tol's kinematics.} .

Considering this sparse sampling of parameter space and the results
from the random forest classifier, it is reasonable to consider that
the best values for $\sigma_{\rm cl}$, $v_{\infty {\rm cl}}$, $P_{\rm
  cl}$ are actually lower/upper/lower bounds, respectively.

@@

\begin{equation}
M_{\rm dyn} = 3 \frac{\sigma ^{2}r}{G} = 3.48\times10^{9}
\left(\frac{\sigma}{100\ \mathrm{km\ s}^{-1}}\right)^2\left(\frac{r}{\mathrm{kpc}}\right)
M_{\odot}, 
\end{equation}
%
Assuming that the \lya\ emission is entirely powered by star formation 
we use the 3D half-light radius $r_s=2.25$ kpc as the typical size
for the HI region, in the multiphase model the clump velocity dispersion
$\sigma_{\rm{cl}}=$\sigmaclump corresponds to a dynamical mass of
$M_{\rm dyn}=2.31\pm0.04 \times 10^{9}$ $M_{\odot}$, which is ten
times the estimated baryonic mass in \tol.
Considering that the results for $\sigma_{\rm{cl}}$ are most probably
a lower bound, the results for the dynamical mass should also be
considered as a lower bound.

@@

A definitive confirmation of our kinematic findings has
to be provided by new observational tests with integral field unit
spectroscopy. 
If our results are confirmed, this would open up a new line of more detailed
enquiries to understand how a compact dwarf galaxy can have the
structural characteristics similar to a dwarf elliptical while being
at the same time isolated and sustaining large starburst.
Such and investigation could provide an evolutionary link between star-forming
and quiescent stages of dwarf galaxy evolution.
